Ornaments as indicators of social changes and cultural practices in northeastern Taiwan before and after European colonial period

### 2019-12-12 analysis/paper/paper.Rmd
make the c14 table work

---
title: "Ornaments as indicators of social changes and cultural practices in northeastern Taiwan before and after European colonial period"
author:
  - Li-Ying Wang
  - Ben Marwick
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
bibliography: references.bib
csl: "../templates/antiquity.csl" # Insert path for the bib-style
abstract: |
  Long-lasting indirect impacts on indigenous peoples in the periphery of colonial control are poorly understood, especially in East Asia. Trade ornaments from Kiwulan (1400-1900 AD) in northeastern Taiwan show the indirect impacts of European colonial activities on local societies. The diversity of ornaments was greater during the period of European presence compared to previous periods, and their spatial distribution was more clustered. This hints at increasing social inequality resulting from a colonial influence. Ornaments give insights into the increasing social inequality stimulated by European colonisation, and show the agency of indigenous people to incorporate ornaments into their social system.
keywords: |
  Historical archaeology; Iron Age; European colonisation; Trade ornaments; Exchange networks; Taiwan; Monte Carlo simulation
highlights: |
  1. Evidence of the quantity, variety, and spatial distribution of trade ornaments in both local cultural and colonial context is provided.
  2. Results of hypothesis testing point to greater amount and diversity of ornament types with non-random, and most highly concentrated spatial clusters during the European presence.
  3. Trade ornament as a prestige good may result in an increase in social inequality due to competition of wealth accumulation between individuals.
  4. Trade ornament can be a good proxy to detect indirect colonial influence on local indigenous populations and inform on social changes relating to trade with colonial groups.
---

@@

```{r, setup, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  cache = TRUE,
  dpi = 300,
  fig.path = "../figures/"
)
# prepare for Antiquity https://antiquity.ac.uk/submit-paper
# should be no more than 13.5cm wide and at a minimum resolution of 300dpi. 
# not exceed 6000 words excluding figures, tables, highlights 
image_width_1_col_inch = 90/25.4 # image Single column	w = 90 mm 
image_width_2_col_inch = 190/25.4 # image Double column w = 190 mm 
```

@@

```{r C14-data, fig.cap= "Summed probability distributions for dates from Kiwulan. The dark line represents the summed probabilities of all radiocarbon ages, and the grey lines in the background are the probabilities of individual ages. Grey rectangles indicate the approximate chronology of the major archaeological components of the deposit. For the upper component, the blue line indicates the start of European presence, while the red line is the Chinese presence. Ages calibrated with the Bchron package (Parnell et al. 2008). "}

@@

# tidy the dataframe 
kw_age_cali_df_tidy <-
  kw_age_cali_df %>% 
  rownames_to_column(var = "Pit-Layer") %>% 
  mutate(`Calibrated date (95% conf. level)`= paste0(`2.5%`," - ", `97.5%`))

@@

```{r C14-table-data}
table_kw_dates <-
  kw_dates %>% 
  mutate(`Above mean sea level (cm)` = paste0(Depth1, " to ", Depth2)) %>%
  mutate(`Pit-Layer` = paste0(Square, "-", Layer)) %>% 
  left_join(kw_age_cali_df_tidy, "Pit-Layer") %>% 
  mutate(ave = (`2.5%` + `97.5%`)/ 2) %>% 
  mutate(ave = ifelse(is.na(ave), 200, ave)) %>%  
  arrange(ave) %>% 
  dplyr::select(`Lab code`, `Pit-Layer`, `Above mean sea level (cm)`, `Date BP`, `Calibrated date (95% conf. level)`, Context) %>% 
  mutate(`Calibrated date (95% conf. level)` = 
           ifelse(is.na(`Calibrated date (95% conf. level)`), "—", `Calibrated date (95% conf. level)`))

knitr::kable(table_kw_dates,
             caption = "Radiocarbon dates on charcoals from Kiwulan (Chen 2007). Calibrated using IntCal13 Atmospheric curve. In the context column, original numbering system from excavation report is used, such M009 means burial No.9 ")
```

>>>>

---
title: "Ornaments as indicators of social changes and cultural practices in northeastern Taiwan before and after European colonial period"
author:
  - Li-Ying Wang
  - Ben Marwick
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::word_document2:
      fig_caption: yes
      #reference_docx: "../templates/template.docx" # Insert path for the DOCX file
bibliography: references.bib
csl: "../templates/antiquity.csl" # Insert path for the bib-style
abstract: |
  Long-lasting indirect impacts on indigenous peoples in the periphery of colonial control are poorly understood, especially in East Asia. Trade ornaments from Kiwulan (1400-1900 AD) in northeastern Taiwan show the indirect impacts of European colonial activities on local societies. The diversity of ornaments was greater during the period of European presence compared to previous periods, and their spatial distribution was more clustered. This hints at increasing social inequality resulting from a colonial influence. Ornaments give insights into the increasing social inequality stimulated by European colonisation, and show the agency of indigenous people to incorporate ornaments into their social system.
keywords: |
  Historical archaeology; Iron Age; European colonisation; Trade ornaments; Exchange networks; Taiwan; Monte Carlo simulation
highlights: |
  1. Evidence of the quantity, variety, and spatial distribution of trade ornaments in both local cultural and colonial context is provided.
  2. Results of hypothesis testing point to greater amount and diversity of ornament types with non-random, and most highly concentrated spatial clusters during the European presence.
  3. Trade ornament as a prestige good may result in an increase in social inequality due to competition of wealth accumulation between individuals.
  4. Trade ornament can be a good proxy to detect indirect colonial influence on local indigenous populations and inform on social changes relating to trade with colonial groups.
---

@@

```{r, setup, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  cache = TRUE,
  dpi = 300,
  fig.path = "../figures/"
)
# prepare for Antiquity https://antiquity.ac.uk/submit-paper
# should be no more than 13.5cm wide and at a minimum resolution of 300dpi. 
# not exceed 6000 words excluding figures, tables, highlights 
image_width_1_col_inch = 90/25.4 # image Single column	w = 90 mm 
image_width_2_col_inch = 190/25.4 # image Double column w = 190 mm

# to make the table work, cf. https://github.com/haozhu233/kableExtra/issues/308
options(kableExtra.auto_format = FALSE)
```

@@

```{r C14-data, fig.cap= "Summed probability distributions for dates from Kiwulan. The dark line represents the summed probabilities of all radiocarbon ages, and the grey lines in the background are the probabilities of individual ages. Grey rectangles indicate the approximate chronology of the major archaeological components of the deposit. For the upper component, the blue line indicates the start of European presence, while the red line is the Chinese presence. Ages calibrated with the Bchron package (Parnell et al. 2008).", cache=TRUE}

@@

# tidy the dataframe 
kw_age_cali_df_tidy <-
  kw_age_cali_df %>% 
  rownames_to_column(var = "Pit-Layer") %>% 
  mutate_at(vars(`2.5%`, `97.5%`), round, 0) %>% 
  mutate(`Calibrated date (95% conf. level)`= paste0(`2.5%`,"-", `97.5%`))

@@

```{r C14-table-data}
table_kw_dates <-
  kw_dates %>% 
  mutate(`Lab code` = str_remove(`Lab code`, " ")) %>% 
  mutate(`Above mean sea level (cm)` = paste0(Depth1, " to ", Depth2)) %>%
  mutate(`Pit-Layer` = paste0(Square, "-", Layer)) %>% 
  left_join(kw_age_cali_df_tidy, "Pit-Layer") %>% 
  mutate(ave = (`2.5%` + `97.5%`)/ 2) %>% 
  mutate(ave = ifelse(is.na(ave), 200, ave)) %>%  
  arrange(ave) %>% 
  dplyr::select(`Lab code`, `Pit-Layer`, 
                `Above mean sea level (cm)`, 
                `Date BP`, 
                `Calibrated date (95% conf. level)`, 
                Context) %>% 
  mutate(`Calibrated date (95% conf. level)` = 
           ifelse(is.na(`Calibrated date (95% conf. level)`), 
                  "", 
                  `Calibrated date (95% conf. level)`))

kableExtra::kable(table_kw_dates,
             caption = "Radiocarbon ages from charcoal samples excavated from Kiwulan (Chen 2007), calibrated using IntCal13 Atmospheric curve. The codes in the context column refer to the excavator's numbering system, cf. ADD CITATION TO THE REPORT HERE")
```
